---
title: "Titanic - Kaggle competition"
output: html_document
date: "2024-02-05"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries

```{r}
library(tidyverse)
```

## Load data

```{r}
titanic_dataset <- 
    read_csv("kaggle/titanic/train.csv")
```

Data splitting

```{r}
titanic_split <- 
    titanic_dataset %>% 
    rsample::initial_split(
        strata = Survived)

titanic_train <- 
    rsample::training(titanic_split)

titanic_validation <- 
    rsample::testing(titanic_split)
```

## EDA

Let's review the data.

We observe there's quite a lot of missing data on Cabin (~80%) and Age (~20%).

```{r}
titanic_train %>%  
    skimr::skim()
```

Let's explore a few theories around who got saved in the 

### Families

An interesting hypothesis to test is whether families were saved together. There's data about the Name that includes the last name, the number of siblings and the number of parents children.

### 





